---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

## Loading and preprocessing the data

```{r data}
##Loading data
data <- read.csv(file = "~/R Working Directory/repdata_data_activity/activity.csv")

##Dealing with Date variable
data$date <- as.POSIXct(data$date, format = "%Y-%m-%d")
```

## Histogram of the total number of steps taken each day

### Creating a new dataset called n_steps

```{r Steps}
library(dplyr)
n_steps <- summarize(group_by(.data = data , date) , Steps = sum(steps))
n_steps
```

### Histogram

```{r Hist_Steps, warning=FALSE}
library(ggplot2)
g <- ggplot(n_steps , aes(x = date , y = Steps))
g+geom_col()

```

## Mean and median number of steps taken each day

```{r Mean_and_median}
Mean <- mean(n_steps$Steps , na.rm = TRUE)
Median <- median(n_steps$Steps , na.rm = TRUE)
cat("Mean = " , Mean , "\n")
cat("Median = " , Median)
```

## Time series plot of the average number of steps taken

### Creating a new dataset called avg_steps

```{r Avg_Steps}
avg_steps <- summarize(group_by(.data = data , interval) , Steps = mean(steps , na.rm = TRUE))
avg_steps
```

### Time Series Plot

```{r Time_series_plot, warning=FALSE}
g <- ggplot(avg_steps , aes(x = interval , y = Steps))
g+geom_line(size = 1 , col = "blue")
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r Max_Steps}
filter(avg_steps , avg_steps$Steps == max(avg_steps$Steps))
```

## Missing values

### total number of missing values in the dataset

```{r n_missing}
sum(is.na(data))
```

### Using average interval for imputing missing data

```{r imputing_na}
data_imputed <- data
for (i in 1:17568) {
  if (is.na(data_imputed$steps[i]))  {
    interval <- data_imputed$interval[i]
    data_imputed$steps[i] <- avg_steps$Steps[avg_steps$interval == interval]
  }
}
```

### Histogram after imputing

```{r Hist_steps2}
n_steps <- summarize(group_by(.data = data_imputed , date) , Steps = sum(steps))
g <- ggplot(n_steps , aes(x = date , y = Steps))
g+geom_col()
```

## Panel plot for weekdays and weekends

### Adding a factor variable to the dataset

```{r factor_var}
data_imputed$day <- ""
for (i in 1:17568) {
  if (weekdays(data_imputed$date[i]) == "Saturday" | weekdays(data_imputed$date[i]) == "Sunday" ) {
    data_imputed$day[i] <- "Weekend"}
  else{
    data_imputed$day[i] <- "Weekday" 
  }
}
data_imputed$day <- factor(x = data_imputed$day)
data_imputed
```

### Creating a new dataset called avg_steps

```{r avg_steps2, message=FALSE, warning=FALSE}
avg_steps <- summarize(group_by(.data = data_imputed , interval , day ) , Steps = mean(steps))
avg_steps
```

### Panel plot

```{r Panel_plot}
g <- ggplot(data = avg_steps , aes(interval , Steps)) + geom_line()
g+ facet_grid(rows = vars(day))
```
